---
title: "6_26_Full_tRNA_Analysis"
output: html_document
date: "2023-06-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading in Data
Creates fullDF, which contains all the EV (summed) and Serum tRNA data
```{r}
library(edgeR)
library(readr)
library(dplyr)
serumDF<-read.csv("totalRNA_serum_JuJu.csv", header = TRUE)
colnames(serumDF) <- c("Genes","Justin_PO_Serum","Justin_1w_Serum","Justin_1m_Serum","Justin_2m_Serum","Julian_PO_Serum","Julian_1w_Serum","Julian_1m_Serum","Julian_2m_Serum")
cleanSerumDF <- serumDF[,-1]
rownames(cleanSerumDF) <- serumDF[,1]

seqInfo_EY<-read_excel("./seqInfo.xlsx")

evDF<- read.csv("totalRNA_JuJu_EV.csv", header = TRUE)
columnHeaders<-colnames(evDF)

cleanEVDF <- evDF[,-1]
rownames(cleanEVDF) <- evDF[,1]

summedEVDF<- cleanEVDF%>%
  rowwise()%>%
  mutate(Justin_PO_EV=sum(c_across(1:3)),
         Justin_1w_EV=sum(c_across(4:6)),
         Justin_1m_EV=sum(c_across(7:9)),
         Justin_2m_EV=sum(c_across(10:12)),
         Julian_PO_EV=sum(c_across(13:15)),
         Julian_1w_EV=sum(c_across(16:18)),
         Julian_1m_EV=sum(c_across(19:21)),
         Julian_2m_EV=sum(c_across(22:24)))%>%
  select(Justin_PO_EV,Justin_1w_EV,Justin_1m_EV,Justin_2m_EV,Julian_PO_EV,Julian_1w_EV,Julian_1m_EV,Julian_2m_EV)
rownames(summedEVDF)<-evDF[,1]

summedEVDF$Genes<- rownames(summedEVDF)
cleanSerumDF$Genes<-rownames(cleanSerumDF)
fullDF<-full_join(cleanSerumDF,summedEVDF,by="Genes")
rownames(fullDF)<-fullDF$Genes
fullDF<-fullDF %>% select(-c("Genes"))
```

## Normalization of Data via filtering lowly expressed genes and filterbyExpr

```{r}
#filtering lowly expressed genes
fullDF_filtered <- fullDF[apply(fullDF == 0, 1, sum) <= 5, ] # Filtering out number of 0 counts
#filterbyExpr
dea<- DGEList(counts=fullDF_filtered,group= seqInfo_EY$Type)
keep <- filterByExpr(dea)
dea <- dea[keep,, keep.lib.sizes=FALSE] 
dea <- calcNormFactors(dea, method = "TMM")
dea_logcpm <- cpm(dea, log = T)
```
Checking via box plot
```{r}
# Normalization and box plot to determine distribution
par(mfrow = c(1,2))
boxplot(as.data.frame(dea_logcpm), 
        las=2, 
        col=as.numeric(dea$samples$group), 
        main = "Non-normalized Data")
boxplot(as.data.frame(dea_logcpm), 
        las=2, 
        col=as.numeric(dea$samples$group), 
        main = "Normalized Data ")

```
Checking via bell curve
```{r}
library(tidyverse)
library(variancePartition)
library(edgeR)
library(limma)
library(RColorBrewer)
library(readxl)
library(pheatmap)
par(mfrow = c(1,2))
plotMDS(dea_logcpm, 
        main = "Dimensions 1 & 2", 
        col = as.numeric(dea$samples$group))
plotMDS(dea_logcpm, 
        main = "Dimensions 3 & 4", 
        col = as.numeric(dea$samples$group), 
        dim.plot = c(3,4))
# Density plot to assess distribution after Filtering
L <- mean(fullDF_filtered$samples$lib.size) * 1e-6
M <- median(fullDF_filtered$samples$lib.size) * 1e-6
query <- as.data.frame(dea_logcpm) %>% 
     pivot_longer(everything(), names_to = "Sample", values_to = "logCPM")
ggplot(query, aes(x = logCPM, fill = Sample)) +
                geom_density(
                        adjust = 1.5,
                        alpha = 0.4
                ) +
                scale_x_continuous(
                        limits = c(-5,24,2),
                        expand = c(0,0),
                        breaks = seq(-5,24,5)
                ) +
                scale_y_continuous(
                        expand = c(0,0)
                ) +
                scale_fill_discrete(
                        breaks = factor(query$Sample)
                ) +
                xlim(-5,24) +
                labs(
                        x = "log-CPM",
                        y = "Density",
                        title = "Merged RNA Density Distribution"
                ) +
                geom_vline(
                     xintercept = log2(10/M + 2/L)
                ) +
                theme(
                        axis.title.x = element_text(margin = margin(10,0,0,0), face = "bold", size = 9),
                        axis.title.y = element_text(margin = margin(10,10,0,0), face = "bold", size = 9),
                        plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
                        panel.background = element_blank(),
                        panel.grid = element_line(color = "gray90"),
                        panel.grid.major = element_line(color = "gray95"),
                        panel.grid.minor = element_blank()
                )
                

```
## PCA
```{r}
seqInfo_EY<-read_excel("./seqInfo.xlsx")
plotpca <- function(df, seq, color, fn, p){ 
  # df = data.frame that acts as the input for the function.
  # seq = sequencing info housing the categories that each sample belongs to.
  # color = colors to be used for PCA plot.
  # fn = filename used for savin the plot.
  # p = path for the file.
  pca_rna <- prcomp(t(df), scale = TRUE, center = TRUE)
  per_var <- 100*(pca_rna$sdev^2/sum(pca_rna$sdev^2))
  d <- data.frame("PC1" = pca_rna$x[,1],
                  "PC2" = pca_rna$x[,2],
                  "PC3" = pca_rna$x[,3],
                  "PC4" = pca_rna$x[,4],
                  "Sheep" = factor(seq$Sheep),
                  "Source" = seq$Type)
  g <- ggplot(d, aes(x = PC1, y = PC3)) +
          geom_point(aes(color = Sheep,
                         shape = Source,
                         size=seqInfo_EY$Time_Size),
                     stroke = 0.8
          ) +
          scale_shape_manual(values = c(15, 17)
          ) +
          scale_color_manual(values = color
          ) +
          labs(
            x = paste0("PC1, VarExp: ", round(per_var[[1]], 1), "%"),
            y = paste0("PC2, VarExp: ", round(per_var[[2]], 1), "%"),
            color = guide_legend(title = "Sheep", size = 18)
          ) +
          theme(
            panel.background = element_blank(),
            panel.grid.minor = element_blank(),
            panel.grid.major = element_blank(),
            text = element_text(size = 18),
            axis.text.x = element_text(size = 16),
            axis.text.y = element_text(size = 16),
            panel.grid = element_line(color = "gray0", linetype = "solid"),
            legend.key = element_blank(),
            legend.text = element_text(size = 16),
            legend.title = element_text(face = "italic")
          ) +
          guides(
            shape = guide_legend(override.aes = list(size = 4)), # Making legend shapes smaller
            color = guide_legend(override.aes = list(size = 4)), # Making legend colors smaller
          ) +
          geom_hline(aes(yintercept = 0), 
                     color = "gray0") +
          geom_vline(aes(xintercept = 0),
                     color = "gray0")
  ggsave(paste0(fn, "PC1_vs_PC2.tiff"), plot = g, device = "tiff", path = p)
}

plotpca(dea_logcpm, seqInfo_EY, color = c("#ec696f", "#53dbf3"), fn = "Justin_vs_Julian_PCA_PC1vsPC3", p = "./6_26_PCA")
```

```{r}
pca_rna <- prcomp(t(dea_logcpm), scale = TRUE, center = TRUE)
loadingscores<-pca_rna$rotation[,1]
gene_scores<-abs(loadingscores)
gene_score_ranked<-sort(gene_scores,decreasing=TRUE)

top_10_genes<-names(gene_score_ranked[1:20])

top_10_genes
```
PCA for Serum
```{r}
plotpca <- function(df, seq, color, fn, p){ 
  # df = data.frame that acts as the input for the function.
  # seq = sequencing info housing the categories that each sample belongs to.
  # color = colors to be used for PCA plot.
  # fn = filename used for savin the plot.
  # p = path for the file.
  pca_rna <- prcomp(t(df), scale = TRUE, center = TRUE)
  per_var <- 100*(pca_rna$sdev^2/sum(pca_rna$sdev^2))
  d <- data.frame("PC1" = pca_rna$x[,1],
                  "PC2" = pca_rna$x[,2],
                  "PC3" = pca_rna$x[,3],
                  "PC4" = pca_rna$x[,4],
                  "Sheep" = factor(seq$Sheep),
                  "Time" = seq$Time)
  g <- ggplot(d, aes(x = PC1, y = PC2)) +
          geom_point(aes(color = Time,
                         shape = Sheep),
                     size=5.5,
                     stroke = 0.8
          ) +
          scale_shape_manual(values = c(15, 17)
          ) +
          scale_color_manual(values = c("#ec696f", "#53dbf1","#538df9","#5b53f9")
          ) +
          labs(
            x = paste0("PC1, VarExp: ", round(per_var[[1]], 1), "%"),
            y = paste0("PC2, VarExp: ", round(per_var[[2]], 1), "%"),
            color = guide_legend(title = "Sheep", size = 18)
          ) +
          theme(
            panel.background = element_blank(),
            panel.grid.minor = element_blank(),
            panel.grid.major = element_blank(),
            text = element_text(size = 18),
            axis.text.x = element_text(size = 16),
            axis.text.y = element_text(size = 16),
            panel.grid = element_line(color = "gray0", linetype = "solid"),
            legend.key = element_blank(),
            legend.text = element_text(size = 16),
            legend.title = element_text(face = "italic")
          ) +
          guides(
            shape = guide_legend(override.aes = list(size = 4)), # Making legend shapes smaller
            color = guide_legend(override.aes = list(size = 4)), # Making legend colors smaller
          ) +
          geom_hline(aes(yintercept = 0), 
                     color = "gray0") +
          geom_vline(aes(xintercept = 0),
                     color = "gray0")
  ggsave(paste0(fn, "PC1_vs_PC2.tiff"), plot = g, device = "tiff", path = p)
}

seqInfo_EY_Serum<-seqInfo_EY[1:8,]

plotpca(dea_logcpm_Serum, seqInfo_EY_Serum, color = c("#ec696f", "#53dbf1","#538df9","#5b53f9"), fn = "Justin_vs_Julian_PCA_Serum", p = "./6_26_PCA")
```

PCA for EV
```{r}
plotpca <- function(df, seq, color, fn, p){ 
  # df = data.frame that acts as the input for the function.
  # seq = sequencing info housing the categories that each sample belongs to.
  # color = colors to be used for PCA plot.
  # fn = filename used for savin the plot.
  # p = path for the file.
  pca_rna <- prcomp(t(df), scale = TRUE, center = TRUE)
  per_var <- 100*(pca_rna$sdev^2/sum(pca_rna$sdev^2))
  d <- data.frame("PC1" = pca_rna$x[,1],
                  "PC2" = pca_rna$x[,2],
                  "PC3" = pca_rna$x[,3],
                  "PC4" = pca_rna$x[,4],
                  "Sheep" = factor(seq$Sheep),
                  "Time" = seq$Time)
  g <- ggplot(d, aes(x = PC1, y = PC2)) +
          geom_point(aes(color = Time,
                         shape = Sheep),
                     size=5.5,
                     stroke = 0.8
          ) +
          scale_shape_manual(values = c(15, 17)
          ) +
          scale_color_manual(values = c("#ec696f", "#53dbf1","#538df9","#5b53f9")
          ) +
          labs(
            x = paste0("PC1, VarExp: ", round(per_var[[1]], 1), "%"),
            y = paste0("PC2, VarExp: ", round(per_var[[2]], 1), "%"),
            color = guide_legend(title = "Sheep", size = 18)
          ) +
          theme(
            panel.background = element_blank(),
            panel.grid.minor = element_blank(),
            panel.grid.major = element_blank(),
            text = element_text(size = 18),
            axis.text.x = element_text(size = 16),
            axis.text.y = element_text(size = 16),
            panel.grid = element_line(color = "gray0", linetype = "solid"),
            legend.key = element_blank(),
            legend.text = element_text(size = 16),
            legend.title = element_text(face = "italic")
          ) +
          guides(
            shape = guide_legend(override.aes = list(size = 4)), # Making legend shapes smaller
            color = guide_legend(override.aes = list(size = 4)), # Making legend colors smaller
          ) +
          geom_hline(aes(yintercept = 0), 
                     color = "gray0") +
          geom_vline(aes(xintercept = 0),
                     color = "gray0")
  ggsave(paste0(fn, "PC1_vs_PC2.tiff"), plot = g, device = "tiff", path = p)
}

seqInfo_EY_EV<-seqInfo_EY[9:16,]

plotpca(dea_logcpm_EV, seqInfo_EY_EV, color = c("#ec696f", "#53dbf1","#538df9","#5b53f9"), fn = "Justin_vs_Julian_PCA_EV", p = "./6_26_PCA")
```



## Heat Map
```{r}


#creating an EV Only Dataset

fullEVDF<-fullDF_filtered %>% select(contains("EV"))

dea_EV<- DGEList(counts=fullEVDF)
keep_EV <- filterByExpr(dea_EV)
dea_EV <- dea_EV[keep_EV,, keep.lib.sizes=FALSE] 
dea_EV <- calcNormFactors(dea_EV, method = "TMM")
dea_logcpm_EV <- cpm(dea_EV, log = T)

#creating an Serum Only Dataset

fullSerumDF<-fullDF_filtered %>% select(contains("Serum"))

dea_Serum<- DGEList(counts=fullSerumDF)
keep_Serum <- filterByExpr(dea_Serum)
dea_Serum <- dea_Serum[keep_Serum,, keep.lib.sizes=FALSE] 
dea_Serum <- calcNormFactors(dea_Serum, method = "TMM")
dea_logcpm_Serum <- cpm(dea_Serum, log = T)
```
Heat Map for Everything
```{r}
# Calculating genes with high variance
var_genes <- apply(dea_logcpm, 1, var)


# 10% Top most variable genes
select_var <- names(sort(var_genes, decreasing = TRUE))[1:(0.01 *length(var_genes))]
#select_var <- names(sort(var_genes, decreasing = TRUE))[1:20]

# converting them to logcpm
hvlcpm <- dea_logcpm[select_var,]

# Drawing heatmap
annot_row = data.frame(Sheep = seqInfo_EY$Sheep,
                       Source = seqInfo_EY$Type,
                       Time=seqInfo_EY$Time)
rownames(annot_row) = colnames(hvlcpm)
annotation_colors <- list(Sheep = c(Julian = "#ec696f", Justin = "#53dbf3"),
                          Source = c(EV = "#9552EE", Serum = "#7be692"),
                          Time=c(PO="#32a6a8",`1W` ="#3269a8",`1M` ="#323ca8",`2M` ="#4e32a8"))

#pdf("./test.tiff", width = 16, height = 16)|
saveThis <- pheatmap(hvlcpm,
         cluster_rows = TRUE,
         cluster_cols = TRUE,
         clustering_distance_cols = "euclidean",
         show_colnames = TRUE,
         show_rownames = TRUE,
         clustering_method = "ward.D2",
         border_color = "gray50",
         scale = "row",
         annotation_col = annot_row,
         annotation_colors = annotation_colors,
         fontsize = 4,
         fontsize_row = 4,
         cellwidth = 4,
         cellheight = 4,
         col = colorRampPalette(c("#32009F", "white", "firebrick3"))(100))

save_pheatmap_pdf <- function(x, filename, width=7, height=7) {
   stopifnot(!missing(x))
   stopifnot(!missing(filename))
   pdf(filename, width=width, height=height)
   grid::grid.newpage()
   grid::grid.draw(x$gtable)
   dev.off()
}
save_pheatmap_pdf(saveThis, "./fullHeat4.pdf")

dev.off()
```


```{r}
# Calculating genes with high variance
var_genes <- apply(dea_logcpm_EV, 1, var)


# 10% Top most variable genes
#select_var <- names(sort(var_genes, decreasing = TRUE))[1:(0.1 *length(var_genes))]
select_var <- names(sort(var_genes, decreasing = TRUE))[1:20]

# converting them to logcpm
hvlcpm <- dea_logcpm_EV[select_var,]

# Drawing heatmap
feedThis = c("Justin","Justin","Justin","Justin","Julian","Julian","Julian","Julian")
         
annot_row = data.frame(Sheep = feedThis)
rownames(annot_row) = colnames(hvlcpm)
annotation_colors <- list(Sheep = c(Julian = "#ec696f", Justin = "#53dbf3"))

#pdf("./test.tiff", width = 16, height = 16)|
saveThis <- pheatmap(hvlcpm,
         cluster_rows = TRUE,
         cluster_cols = TRUE,
         clustering_distance_cols = "euclidean",
         show_colnames = FALSE,
         show_rownames = TRUE,
         clustering_method = "ward.D2",
         border_color = "gray50",
         scale = "row",
         annotation_col = annot_row,
         annotation_colors = annotation_colors,
         fontsize = 5,
         fontsize_row = 5,
         cellwidth = 5,
         cellheight = 5,
         col = colorRampPalette(c("#32009F", "white", "firebrick3"))(100))

save_pheatmap_pdf <- function(x, filename, width=7, height=7) {
   stopifnot(!missing(x))
   stopifnot(!missing(filename))
   pdf(filename, width=width, height=height)
   grid::grid.newpage()
   grid::grid.draw(x$gtable)
   dev.off()
}
save_pheatmap_pdf(saveThis, "./EVs.pdf")

#dev.off()
```



```{r}
# Calculating genes with high variance
var_genes <- apply(dea_logcpm_Serum, 1, var)


# 10% Top most variable genes
#select_var <- names(sort(var_genes, decreasing = TRUE))[1:(0.1 *length(var_genes))]
select_var <- names(sort(var_genes, decreasing = TRUE))[1:20]

# converting them to logcpm
hvlcpm <- dea_logcpm_Serum[select_var,]

# Drawing heatmap
feedThis = c("Justin","Justin","Justin","Justin","Julian","Julian","Julian","Julian")
         
annot_row = data.frame(Sheep = feedThis)
rownames(annot_row) = colnames(hvlcpm)
annotation_colors <- list(Sheep = c(Julian = "#ec696f", Justin = "#53dbf3"))

#pdf("./test.tiff", width = 16, height = 16)|
saveThis <- pheatmap(hvlcpm,
         cluster_rows = TRUE,
         cluster_cols = TRUE,
         clustering_distance_cols = "euclidean",
         show_colnames = FALSE,
         show_rownames = TRUE,
         clustering_method = "ward.D2",
         border_color = "gray50",
         scale = "row",
         annotation_col = annot_row,
         annotation_colors = annotation_colors,
         fontsize = 5,
         fontsize_row = 5,
         cellwidth = 5,
         cellheight = 5,
         col = colorRampPalette(c("#32009F", "white", "firebrick3"))(100))

save_pheatmap_pdf <- function(x, filename, width=7, height=7) {
   stopifnot(!missing(x))
   stopifnot(!missing(filename))
   pdf(filename, width=width, height=height)
   grid::grid.newpage()
   grid::grid.draw(x$gtable)
   dev.off()
}
save_pheatmap_pdf(saveThis, "./Serum.pdf")

#dev.off()
```

## Analysis with Dream




#### DGE Analysis of Justin Serum
```{r}
seqInfo_EY_JustinSerum <- subset(seqInfo_EY, Sheep %in% c("Justin"))[1:4,]
dfJustinSerum<-fullDF_filtered%>% select(contains("Justin"))
dfJustinSerum <- dfJustinSerum[,1:4]
Operation<-factor(seqInfo_EY_JustinSerum$Operation, levels = c('Pre', 'Post'))

designEY_JustinSerum<-model.matrix(~ Operation)


form <- ~ Operation #expression is influenced by both operation and time

param <- SnowParam(4, "SOCK", progressbar = TRUE)

vobjDreamEY <- voomWithDreamWeights(dfJustinSerum, form, seqInfo_EY_JustinSerum, BPPARAM = param)
fitmmEY <- dream(vobjDreamEY, form, seqInfo_EY_JustinSerum)
fitmmEY <- eBayes(fitmmEY)

finalEY <- topTable(fitmmEY, coef = 2, sort.by = "logFC", number = Inf) %>%
     mutate(log10P = -log10(adj.P.Val), .after = adj.P.Val)
write.csv(finalEY, "./JustinSerum.csv")

```

##### Further Analysis of Justin Serum Week by Week
https://rstudio-pubs-static.s3.amazonaws.com/525119_64c1fe6e1a514b89a1ef26d23bf4aae3.html 
Via NOISeq
```{r}
library(NOISeq)
dfJustinSerum1W <- dfJustinSerum[,c(1,2)]
SampleName <- c("Justin_PO_Serum", "Justin_1w_Serum")
Time <- c("Pre-Op", "1W")

factorEY=data.frame(SampleName,Time)
rownames(factorEY)=factorEY[,1]
#factorEY2 <- factorEY[,-1]
#rownames(factorEY2) <- factorEY[,1]

inputEY=readData(dfJustinSerum1W,factorEY)
NOISEQ1W= noiseq(inputEY, k = 0.5, norm = "rpkm", 
replicates = "no", 
factor="Time", conditions=c("Justin_PO_Serum","Justin_1w_Serum"), pnr = 0.2, nss = 5, v = 0.02, lc = 0)
```
```{r}
NOISEQ1W.deg = degenes(NOISEQ1W, q = 0.8, M = NULL)
```

```{r}
par(mfrow = c(1, 2))

#Plot expression with DEGs highlighted in red
DE.plot(NOISEQ1W, q = 0.8, graphic = "expr", log.scale = TRUE)

```
Pre-Op vs 1M
```{r}
dfJustinSerum1M <- dfJustinSerum[,c(1,3)]
SampleName <- c("Justin_PO_Serum", "Justin_1m_Serum")
Time <- c("Pre-Op", "1M")

factorEY=data.frame(SampleName,Time)
rownames(factorEY)=factorEY[,1]
#factorEY2 <- factorEY[,-1]
#rownames(factorEY2) <- factorEY[,1]

inputEY=readData(dfJustinSerum1M,factorEY)
NOISEQ1M= noiseq(inputEY, k = 0.5, norm = "rpkm", 
replicates = "no", 
factor="Time", conditions=c("Justin_PO_Serum","Justin_1m_Serum"), pnr = 0.2, nss = 5, v = 0.02, lc = 0)

NOISEQ1M.deg = degenes(NOISEQ1M, q = 0.8, M = NULL)
par(mfrow = c(1, 2))

#Plot expression with DEGs highlighted in red
DE.plot(NOISEQ1M, q = 0.8, graphic = "expr", log.scale = TRUE)

```
Pre-Op vs 2M
```{r}
dfJustinSerum2M <- dfJustinSerum[,c(1,4)]
SampleName <- c("Justin_PO_Serum", "Justin_2m_Serum")
Time <- c("Pre-Op", "2M")

factorEY=data.frame(SampleName,Time)
rownames(factorEY)=factorEY[,1]
#factorEY2 <- factorEY[,-1]
#rownames(factorEY2) <- factorEY[,1]

inputEY=readData(dfJustinSerum2M,factorEY)
NOISEQ2M= noiseq(inputEY, k = 0.5, norm = "rpkm", 
replicates = "no", 
factor="Time", conditions=c("Justin_PO_Serum","Justin_2m_Serum"), pnr = 0.2, nss = 5, v = 0.02, lc = 0)

NOISEQ2M.deg = degenes(NOISEQ2M, q = 0.8, M = NULL)
par(mfrow = c(1, 2))

#Plot expression with DEGs highlighted in red
DE.plot(NOISEQ2M, q = 0.8, graphic = "expr", log.scale = TRUE)

```
#### DGE Analyis of Justin EV 

```{r}
seqInfo_EY_JustinEV <- subset(seqInfo_EY, Sheep %in% c("Justin"))[5:8,]
dfJustinEV<-fullDF_filtered%>% select(contains("Justin"))
dfJustinEV <- dfJustinEV[,5:8]
Operation<-factor(seqInfo_EY_JulianSerum$Operation, levels = c('Pre', 'Post'))

designEY_JustinEV<-model.matrix(~ Operation)

form <- ~ Operation #expression is influenced by both operation and time

param <- SnowParam(4, "SOCK", progressbar = TRUE)

vobjDreamEY <- voomWithDreamWeights(dfJustinEV, form, seqInfo_EY_JustinEV, BPPARAM = param)
fitmmEY <- dream(vobjDreamEY, form, seqInfo_EY_JustinEV)
fitmmEY <- eBayes(fitmmEY)

finalEY <- topTable(fitmmEY, coef = 2, sort.by = "logFC", number = Inf) %>%
     mutate(log10P = -log10(adj.P.Val), .after = adj.P.Val)
write.csv(finalEY, "./JustinEV.csv")

```

#### Julian EV DGE
```{r}
seqInfo_EY_JulianEV <- subset(seqInfo_EY, Sheep %in% c("Julian"))[5:8,]
dfJulianEV<-fullDF_filtered%>% select(contains("Julian"))
dfJulianEV <- dfJulianEV[,5:8]
Operation<-factor(seqInfo_EY_JulianSerum$Operation, levels = c('Pre', 'Post'))

designEY_JulianEV<-model.matrix(~ Operation)

form <- ~ Operation #expression is influenced by both operation and time

param <- SnowParam(4, "SOCK", progressbar = TRUE)

vobjDreamEY <- voomWithDreamWeights(dfJulianEV, form, seqInfo_EY_JulianEV, BPPARAM = param)
fitmmEY <- dream(vobjDreamEY, form, seqInfo_EY_JulianEV)
fitmmEY <- eBayes(fitmmEY)

finalEY <- topTable(fitmmEY, coef = 2, sort.by = "logFC", number = Inf) %>%
     mutate(log10P = -log10(adj.P.Val), .after = adj.P.Val)
write.csv(finalEY, "./JulianEV.csv")

```
#### DGE Analysis of Julian Serum
```{r}
seqInfo_EY_JulianSerum <- subset(seqInfo_EY, Sheep %in% c("Julian"))[1:4,]
dfJulianSerum<-fullDF_filtered%>% select(contains("Julian"))
dfJulianSerum <- dfJulianSerum[,1:4]
Operation<-factor(seqInfo_EY_JulianSerum$Operation, levels = c('Pre', 'Post'))

designEY_JulianSerum<-model.matrix(~ Operation)

form <- ~ Operation #expression is influenced by both operation and time

param <- SnowParam(4, "SOCK", progressbar = TRUE)

vobjDreamEY <- voomWithDreamWeights(dfJulianSerum, form, seqInfo_EY_JulianSerum, BPPARAM = param)
fitmmEY <- dream(vobjDreamEY, form, seqInfo_EY_JulianSerum)
fitmmEY <- eBayes(fitmmEY)

finalEY <- topTable(fitmmEY, coef = 2, sort.by = "logFC", number = Inf) %>%
     mutate(log10P = -log10(adj.P.Val), .after = adj.P.Val)
write.csv(finalEY, "./JulianSerum.csv")

```




#### bad below


```{r dream warning = FALSE, message = FALSE}
# Setting up design matrix
design <- model.matrix(~ seqInfo_EY$Type, seqInfo_EY)
colnames(design) <- c("(Intercept)", "Source EV")
# Dream
param <- SnowParam(4, "SOCK", progressbar = TRUE)
form <- ~ Type + (1|Sheep)
vobjDream <- voomWithDreamWeights(fullDF, form, seqInfo_EY, BPPARAM = param)
fitmm <- dream(vobjDream, form, seqInfo_EY)
fitmm <- eBayes(fitmm)
head(fitmm$design, 3)
# Here, the log fold change is relative to the EV. E.g. logFC = 10 means that miRNA is 20 times more expressed in the EV when compared to the cell. 
```
```{r}
final <- topTable(fitmm, coef = "TypeSerum", sort.by = "logFC", number = Inf) %>%
     mutate(log10P = -log10(adj.P.Val), .after = adj.P.Val)
write.csv(final, "./test.csv")
rm(list = setdiff(ls(), c("m_rna_logcpm", "final", "e_rna_logcpm", "c_rna_logcpm", "fitmm")))
```

Evan FIrst try below

```{r}
seqInfo_EY_Justin <- subset(seqInfo_EY, Sheep %in% c("Justin"))
dfJustin<-fullDF_filtered%>% select(contains("Justin"))

designEY_Justin<-model.matrix(~ seqInfo_EY_Justin$Operation)

form <- ~ Operation + (1|Type)  #these are the variables that may change 

param <- SnowParam(4, "SOCK", progressbar = TRUE)

vobjDreamEY <- voomWithDreamWeights(dfJustin, form, seqInfo_EY_Justin, BPPARAM = param)
fitmmEY <- dream(vobjDreamEY, form, seqInfo_EY_Justin)
fitmmEY <- eBayes(fitmmEY)

finalEY <- topTable(fitmmEY, coef = "OperationPre", sort.by = "logFC", number = Inf) %>%
     mutate(log10P = -log10(adj.P.Val), .after = adj.P.Val)
write.csv(finalEY, "./Justin_All22.csv")

```

```{r}
seqInfo_EY_Julian <- subset(seqInfo_EY, Sheep %in% c("Julian"))
dfJulian<-fullDF_filtered%>% select(contains("Julian"))

designEY_Julian<-model.matrix(~ seqInfo_EY_Julian$Operation)

form <- ~ Operation + (1|Type)  #these are the variables that may change 

param <- SnowParam(4, "SOCK", progressbar = TRUE)

vobjDreamEY <- voomWithDreamWeights(dfJulian, form, seqInfo_EY_Julian, BPPARAM = param)
fitmmEY <- dream(vobjDreamEY, form, seqInfo_EY_Julian)
fitmmEY <- eBayes(fitmmEY)

finalEY <- topTable(fitmmEY, coef = "OperationPre", sort.by = "logFC", number = Inf) %>%
     mutate(log10P = -log10(adj.P.Val), .after = adj.P.Val)
write.csv(finalEY, "./Julian_All22.csv")

```